---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# wildlifestrikes

<!-- badges: start -->
<!-- badges: end -->

## Overview

 "wildlifestrikes" is a package which analyses the wildlife strikes to aircrafts in the United States from 1991 to 2018. 
 
## Installation 

You can install the released version of wildlifestrikes from [CRAN](https://CRAN.R-project.org) with:

```{r eval=FALSE}
install.packages("wildlifestrikes")
```

You can install the development version from GitHub with:

```{r eval=FALSE}
# install.packages("devtools")
#devtools::install_github("")
```

## Purposes of this package 

The goals of the package are to:

- Inform users of the conditions and factors that affect the frequency and occurrence of wildlife strikes on aircrafts 
- Analyse the number of cases of wildlife strikes on aircrafts in the user selected states 
- Analyse the impact of wildlife strikes on airlines in terms of number of cases, damage levels and repair costs
- Calculate the correlation and p-value between selected variable and repair cost incurred due to wildlife strikes 
- Provide an easy to use and clear platform to digest and understand the data using Shiny


## Usage 

```{r}
library(wildlifestrikes)
```

The below examples use the built-in dataset `wildlife_strikes` and `USmap`. 

The package contains the following functions:

1) The function, `count_frequency()` calculates the frequency of the observations for a particular variable in the `wildlife_strikes` data set in a tibble format while `count_frequency_plot()` shows the frequency of observations in a bar plot. It is helpful in identifying which condition in the variable is associated with higher occurrence of wildlife strikes to aircrafts. This function has been used in the Shiny App as well. 

```{r }
count_frequency(sky)
```


```{r }
count_frequency_plot(sky)
```


2) The function, `cases_statemap()` presents the number of wildlife strikes reported in the selected states from 1991 to 2018 in  a choropleth map. Alternatively, run `cases_state()` to show the observations for the selected states in a tibble format. This function has been used in the Shiny App as well. 

```{r }
cases_statemap(c("arizona", "california", "florida", "colorado", "texas"))
```

```{r}
cases_state(c("arizona", "california", "florida", "colorado", "texas"))
```


3) The function, `cases_airline()`returns a message to you on the number of attacks incurred by your selected airline in the your selected year. This function has been used in the Shiny App as well. 

```{r }
cases_airline("AMERICAN AIRLINES", 2018)
```

4) You can get further information on the number of attacks by your selected airline in your selected year by damage levels by running the function `damages_airline()`. This function has been used in the Shiny App as well. 

```{r}
damages_airline("AMERICAN AIRLINES", 2011)
```

5) The function `hist.R()`, builds the histogram which shows the distribution of repair costs incurred for the selected airline from 1991 to 2018 due to wildlife strikes. This function has been used in the Shiny App as well. 

```{r}
histPlot("AMERICAN AIRLINES", 10)
```

6) The function **`calculate_corr()`** computes the correlation and p-value between that selected variable and repair costs incurred on airline due to wildlife strikes. This is helpful in finding the relationship between the variable and repair costs incurred on airline due to wildlife strikes.

```{r}
calculate_corr(var = height)
```



7) There are 2 functions built solely to support the Shiny
  - `ui_inputs.R` creates user interface inputs for the Shiny application
  - `theme_plot.R` creates a custom theme for the bar plots and maps

## Launch Shiny App 

To launch the shiny app, run the following:

```{r eval=FALSE}
run_app()
```

