---
title: "wildlifestrikes_aircrafts_USA"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{wildlifestrikes_aircrafts_USA}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Overview 

The number of wildlife strikes reported per year to the Federal Aviation Administration (FAA) increased steadily from about 1,800 in 1990 to 16,000 in 2018. Factors contributing to the increase include expansion of wildlife populations, increases in number of aircraft movements, a trend toward faster and quieter aircraft, and outreach to the aviation community.  strikes, there has been greater emphasis on wildlife strike hazard research and airfield wildlife management. As a result,  aircraft collisions with birds and other wildlife (wildlife strikes) have become a growing concern for aviation safety.  Globally, wildlife strikes killed more than 293 people and
destroyed over 271 aircraft from 1988-2020.

This vignette shows you an analysis of wildlife strikes to aircrafts in the United States from 1991 to 2018, using the `wildlifestrikes` package.

Dataset is obtained from [tidytuesday](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-07-23). The dataset contains obtained using this link reported on wildlife impacts from the big 4 airlines in the United States, namely American Airlines, Delta, Southwest, and United as they account for around 70% of passengers in the USA.


# Datasets

The clean raw datasets can be retrieved with theses functions embedded in the package:

- `wildlife_strikes` : Contains wildlife strikes cases on a given date across the United States from 1991 to 2018.

- `USmap`: Contains the geographical coordinates for each state, primarily used for relational joins. 

The dataset is combined with the U.S map data to provide an interactive map and other data visualisations to allow users to explore wildlife strike cases across the U.S.

# Purposes of this package 

<br>

The goals of the package are to:

- Inform users of the conditions and factors that affect the frequency and occurrence of wildlife strikes on aircrafts 
- Analyse the number of cases of wildlife strikes on aircrafts in the user selected states 
- Analyse the impact of wildlife strikes on airlines in terms of number of cases, damage levels and repair costs
- Calculate the correlation and p-value between selected variable and repair cost incurred due to wildlife strikes 
- Provide an easy to use and clear platform to digest and understand the data using Shiny

<br>


# Load package 

You are required to load/download the R package by launching the `library()` function. 

```{r setup}
# devtools::install_github("etc5523-2022/rpkg-Minminsoh")

library(wildlifestrikes) 
```


As mentioned previously, you could use the functions individually which are demonstrated in the following sections. 

Alternatively, some of these functions have been embedded in the Shiny application as well, which provides an easy to use and clear platform to digest and understand the data. You could launch the shiny app with a single line of code: 

```{r eval=FALSE}
run_app()
```

## Goal 1: Inform users of the conditions and factors that affect the frequency and occurrence of wildlife strikes on aircrafts 

**`count_frequency()`** is the function that calculates the frequency of occurrence for a particular variable in the `wildlife_strikes` dataset in a **dataframe** format. It is helpful in identifying which condition in the variable is associated with higher occurence of wildlife strikes to aircrafts. 

A common way to get a better understanding is to visualise it. This package also provides an alternative function **`count_frequency_plot`** which presents the same results in a bar plot format. 

More details about the usage of these functions can be found by using the `help(count_frequency)` and `help(count_frequency_plot)` function.

### Arguments

Both functions take 1 argument, being `var`. You only need to provide the variable, which will then calculates the frequency of the variable for you.

```{r echo=FALSE}
df <- data.frame(Arguments = "`var`",
                Description = "the name of variable or the the name of the column in the dataset")
                                 
knitr::kable(df)
```

### Example

For example, `sky` variable is used, which then computes the frequency of wildlife strikes to aircrafts in different sky conditions.
In this case, wildlife strikes tend to occur the most when there is no cloud. 

```{r}
count_frequency(sky)
```

```{r}
count_frequency_plot(sky)
```

## Goal 2: Analyse the number of cases of wildlife strikes on aircrafts in the user selected states 

By running the function **`cases_statemap(input_state)`**, you would be able to visualise the number of cases of wildlife strikes on aircrafts in the states selected in a choropleth map. The Choropleth map is implemented with a sequential colour scale to establish the geographical location of the states with the highest total wildlifestrike cases from 1991 to 2018.

Alternatively, function **`cases_state(input_state)`** presents the same information in a dataframe format.

More details about the usage of these functions can be found by using the `help(cases_statemap)` and `help(cases_state)` function.

### Arguments

Both functions take 1 argument, being `input_state`. 

```{r echo=FALSE}
df <- data.frame(Arguments = "`input_state`",
                Description = "Name of the states in the dataset")
                                 
knitr::kable(df)

```

### Example

For example, if `input_state` consists of "arizona", "california", "florida", "colorado", "texas", this will return a choropleth map showing the number of total wildlifestrike cases from 1991 to 2018. California has the highest cases of wildlife strikes, 34 cases from 1991 to 2018.

```{r}
cases_statemap(c("arizona", "california", "florida", "colorado", "texas"))
```

```{r}
cases_state(c("arizona", "california", "florida", "colorado", "texas"))
```

## Goal 3: Analyse the impact of wildlife strikes on airlines in terms of number of cases, damage levels and repair costs

There are 3 functions in this section which analyses the impact of the wildlife strikes on airlines. The first function is analysing the number of cases, second function to further categoring the cases to damage levels and the last function analyses the distribution of repair costs incurred during the period. 

### Impact on Airlines : Number of cases 

Function **`cases_airline()`** returns a message to you on the number of attacks incurred by your selected airline in the your selected year.

More details about the usage of this function can be found by using the `help(cases_airline)` function.

#### Arguments

This function takes 2 arguments, being `input_operator` and `input_year`.  

```{r echo=FALSE}

df <- data.frame(Arguments = c("`input_operator`", "`input_year`"),
                Description = c("Name of the airlines in the dataset", "Year in the dataset, ranging from 1990 to 2018"))
                                 
knitr::kable(df)
```

#### Examples

```{r}
cases_airline("AMERICAN AIRLINES", 2018)
```

### Impact on Airlines : Damage levels

Run function **`damages_airline()`** which computes the number of wildlife attacks incurred by the selected airline in the selected year by damages level.

More details about the usage of this function can be found by using the `help(damages_airline)` function.

#### Arguments

This function takes 2 arguments, being `input_operator` and `input_year`.  

```{r echo=FALSE}

df <- data.frame(Arguments = c("`input_operator`", "`input_year`"),
                Description = c("Name of the airlines in the dataset", "Year in the dataset, ranging from 1990 to 2018"))
                                 
knitr::kable(df)
```

#### Examples

For example, if you are interested in the damage levels of American Airlines in 2011, you can set the `input_operator` to be "AMERICAN AIRLINES" and `input_year` to be 2011. In this case, there are 3 cases where damage level is minor, 2 cases where damage level is substantial and 1 case where damage level is uncertain.


```{r}
damages_airline("AMERICAN AIRLINES", 2011)
```

### Impact on Airlines : Repair costs

By running the function **`histPlot()`**, this builds the histogram which shows the distribution of repair cost incurred for the selected airline from 1991 to 2018 due to wildlife strikes, where data is obtained from the `wildlife_strikes` dataset. 

More details about the usage of this function can be found by using the `help(histPlot)` function.

#### Arguments

This function takes 2 arguments, being `input_operator` and `input_bins`.  

```{r echo=FALSE}

df <- data.frame(Arguments = c("`input_operator`", "`input_bins`"),
                Description = c("Name of the airlines in the dataset", "Number of the bins for histogram"))
                                 
knitr::kable(df)
```

#### Example 

For example, if you are interested in the distribution of repair costs incurred by American Airlines due to wildlife strikes from 1990 to 2018, you can set the `input_operator` to be "AMERICAN AIRLINES". You can decide the number of bins to be set in the histogram, for example setting 10 bins in this case.The distribution is right-skewed, there are a few expensive repairs incurred by American Airlines.

```{r}
histPlot("AMERICAN AIRLINES", 10)
```

## Goal 4: Calculate the correlation and p-value between selected variable and repair cost incurred due to wildlife strikes

**`calculate_corr()`** takes 1  variable from `wildlife_strikes` dataset and computes the correlation and p-value between that selected variable and repair costs incurred on airline due to wildlife strikes. This is helpful in finding the relationship between the variable and repair costs incurred on airline due to wildlife strikes.

More details about the usage of this function can be found by using the `help(calculate_corr)` function.

### Argument

This function takes 1 argument, being `var`.  

```{r echo=FALSE}

df <- data.frame(Arguments = "`var`",
                Description = "Name of variable in dataset")
                                 
knitr::kable(df)
```

### Example

For example, if you are interested if height of the plane is associated with the repair costs incurred, the input `var` would be height. In this case, it appears that height has a pretty weak negative relationship with repair costs. 

```{r}
calculate_corr(var = height)
```






